---
title: "TAI Geochemistry Analysis"
author: "AMP, KH, DD"
date: '2022-08-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Create Function to download data from google drive

```{r Make Functions}
## Create a function to read in data
read_data <- function(data){
  # read in data
  read_csv(file = data) %>%
  mutate(transect_location = tolower(transect_location)) 
}

```
## Get Data from GDrive

We need to import data from Google Drive folder.
The list is:
1. TC/TN
2. Bulk Density 
3. GWC
4. LOI
5. pH (soil)


```{r get data}

require(pacman)
pacman::p_load(tidyverse, # keep things tidy
               janitor, # useful for simplifying column names
               googlesheets4, # read_sheet 
               googledrive,# drive_ functions
               readr,# read in file types
               readxl) # read in excel files

## Set GDrive URL raw data files
directory = "https://drive.google.com/drive/folders/1y5a3l0IFp29ofLZC6LGy5E_WSgL5b3qy"
KitMetadata = "https://drive.google.com/file/d/1auYnbc16eg6Dg6AzUJ_DNP5UOd2H-7Kl/view?usp=sharing"

## Create a list of files to download
files <- drive_ls(directory) %>% 
  filter(grepl("_soil_", name))

## Download files to local (don't worry, we'll delete em in a sec)
lapply(files$name, drive_download, overwrite = TRUE)

## Read in datasets
soils_data_all <- files$name %>% 
  map(read_data) %>% 
  reduce(full_join, by= c("campaign", "kit_id", "transect_location")) %>%
  mutate(kit_id = str_trim(kit_id, side = c("both", "left", "right")),
         transect_location = str_trim(transect_location, side = c("both", "left", "right")),
         campaign = str_trim(campaign, side = c("both", "left", "right"))) %>%
  bind_cols() 

#Metadata 
KitMD = googledrive::drive_download(KitMetadata, overwrite = TRUE)
meta.raw = read_csv(file = KitMD$name)

## Clean up local (delete downloaded files)
file.remove(KitMD$name)
file.remove(c(files$name))

#read in additional metadata from Donnie (need to get more details from him)
Donnie_soils_type <- read_csv("data/220713_soil_types_from_donnie.csv", show_col_types = FALSE)

```
#Get rid of things that don't exist 
```{r missing}
metadata_directory = "https://drive.google.com/drive/folders/1IQUq_sD-Jama7ajaZl1zW_9zlWfyCohn"

updated_metadata_directory = "https://drive.google.com/drive/folders/1xy9nsTpbiSk-KXrcDEeFQ7gg4350DuH-"



## Next, list all files in the directory, filter out the Kit Level file and snag the file name
metadata_file <- drive_ls(updated_metadata_directory) %>%
  filter(grepl("kitlevel", name)) %>%
  pull(name)


metadata_file_2 <- drive_ls(updated_metadata_directory) %>%
  filter(grepl("collectedlevel", name)) %>%
  pull(name)

## Now download that file to your local (an annoying googledrive requirement)
drive_download(metadata_file, overwrite = T)
drive_download(metadata_file_2, overwrite = T)



## Finally, read those data in and format so we can set up a list of all samples received
metadata_collected <- read_csv(metadata_file) %>%
  select(kit_id, samples_collected) %>%
   mutate(sediment = ifelse(str_detect(samples_collected, "Sediment"), T, F),
          wetland = ifelse(str_detect(samples_collected, "Wetland"), T, F),
          transition = ifelse(str_detect(samples_collected, "Transition"), T, F),
          upland = ifelse(str_detect(samples_collected, "Upland"), T, F)) %>%
  pivot_longer(cols = c(sediment, wetland, transition, upland),
               names_to = "transect_location", values_to = "collected")

metadata_collected_2 <- read_csv(metadata_file_2)

# %>%
#   pivot_longer(cols= -kit_id, names_pattern = "[:alpha:](?=_)", names_to = "transect_location")

#saveRDS(metadata_collected, "./soils_data_collected_samples_list.rds")

```

##Clean soils metadata
Do we know why there are NAs in these metrics for certain kits?
e.g. K012 Transition, Upland, K026 etc 

```{r soil metadata, echo=FALSE}
collection_metadata_soiltypes_raw_1 <-  Donnie_soils_type %>%
   pivot_longer(cols = c(wetland_soil_horizon, transition_soil_horizon, upland_soil_horizon),
               names_to = 'transect_location') %>% 
  select(kit_id, transect_location, value) %>%
  rename(soil_horizon = value) %>%
  mutate(transect_location = stringr::str_replace(transect_location,"_soil_horizon", "")) 
    # mutate(transect_location = stringr::str_replace(transect_location,"upland","Upland"))  %>%
    # mutate(transect_location = stringr::str_replace(transect_location,"transition","Transition"))  %>% 
    # mutate(transect_location = stringr::str_replace(transect_location,"wetland","Wetland"))  

collection_metadata_soiltypes_raw_2 <-  Donnie_soils_type %>%
   pivot_longer(cols = c(`wt_Percent clay`, `tr_Percent clay`, `up_Percent clay`),
               names_to = 'transect_location') %>% 
  select(kit_id, transect_location, value) %>%
  rename(percent_clay = value) %>%
  mutate(transect_location = stringr::str_replace(transect_location,"_Percent clay", "")) %>%
    mutate(transect_location = stringr::str_replace(transect_location,"up","upland"))  %>%
    mutate(transect_location = stringr::str_replace(transect_location,"tr","transition"))  %>% 
    mutate(transect_location = stringr::str_replace(transect_location,"wt","wetland"))  

collection_metadata_soiltypes_raw_3 <-  Donnie_soils_type %>%
   pivot_longer(cols = c(`wt_Percent sand`, `tr_Percent sand`, `up_Percent sand`),
               names_to = 'transect_location') %>% 
  select(kit_id, transect_location, value) %>%
  rename(percent_sand = value) %>%
  mutate(transect_location = stringr::str_replace(transect_location,"_Percent sand", "")) %>%
    mutate(transect_location = stringr::str_replace(transect_location,"up","upland"))  %>%
    mutate(transect_location = stringr::str_replace(transect_location,"tr","transition"))  %>% 
    mutate(transect_location = stringr::str_replace(transect_location,"wt","wetland"))  

collection_metadata_soiltypes_a <- full_join(collection_metadata_soiltypes_raw_1, collection_metadata_soiltypes_raw_2, by=c("kit_id","transect_location"))

collection_metadata_soiltypes <- full_join(collection_metadata_soiltypes_a, collection_metadata_soiltypes_raw_3, by=c("kit_id","transect_location"))

collection_metadata_soiltypes <- collection_metadata_soiltypes %>%
 mutate(campaign= "EC1") %>%
 drop_na(-kit_id, -transect_location, -campaign) #need to drop rows where all 3 variables are NA

```

```{r visual metrics, echo=FALSE}
visual = googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1Swr2Qx-mBXrhRPIpGf-YehSEC6zOwTPKB3u2K9s2Z7A/edit#gid=1593904722")

visual_clean = visual %>%
  mutate(kit_id = stringr::str_extract(Sample,"K[:digit:]{3}" ),
         transect_location = stringr::str_extract(Sample,"upland|transition|wetland|sediment" ),
         campaign = "EC1") %>%
    # mutate(transect_location = stringr::str_replace(transect_location,"upland","Upland"))  %>%
    # mutate(transect_location = stringr::str_replace(transect_location,"transition","Transition"))  %>% 
    # mutate(transect_location = stringr::str_replace(transect_location,"wetland","Wetland")) %>%
    # mutate(transect_location = stringr::str_replace(transect_location,"sediment","Sediment")) %>%
  rename(sample_weight_g = `Sample weight (from tare) (g)`,
         plant_litter= `Plant litter`,
         visible_minerals = `Visible Minerals`,
         visible_iron_ox = `Presence of Iron Oxidation`) %>%
 select(campaign, kit_id, transect_location, sample_weight_g, plant_litter, visible_minerals, visible_iron_ox)
```
               
##Merge and save
```{r merge and save, echo=FALSE}

meta1 = left_join(metadata_collected, meta.raw, by=c("kit_id", "samples_collected"))

meta = meta1 %>% filter(collected == TRUE) %>%
  mutate(campaign = "EC1")

soils_data_all_wmeta_a <- full_join(soils_data_all, meta, by=c("kit_id", "transect_location", "campaign"))

soils_data_all_wmeta1 <- full_join(soils_data_all_wmeta_a, collection_metadata_soiltypes, by=c("kit_id","transect_location","campaign"))

#something weird happening with kit 057

#also we have a bunch of things that were theoretically not collected... 
#and visual metrics where we have no other data...

soils_data_all_wmeta_vis <- full_join(soils_data_all_wmeta1, visual_clean, by=c("kit_id", "transect_location", "campaign"))
  
# Save as rds
saveRDS(soils_data_all, "./soils_data_merged.rds")
saveRDS(soils_data_all_wmeta_vis, "./soils_data_merged_withmeta.rds")
```

LOI missing datasets 
```{r LOI sass, echo=FALSE}

sample_mass <- visual_clean %>%
  select(kit_id, transect_location, sample_weight_g)

LOI_data <- soils_data_all %>%
  select(kit_id, transect_location, loi_perc)

LOI_TC_data <- soils_data_all %>%
  select(kit_id, transect_location, loi_perc, tc_perc) %>%
  mutate(LOI_TC_ratio = loi_perc/tc_perc,
         STC = (0.441 * loi_perc)+56 )


LOI_missing <- full_join(sample_mass, LOI_data, by= c("kit_id", "transect_location"))

LOI_missing1 <- LOI_missing %>% filter(is.na(loi_perc))

write_csv(LOI_missing1, "./LOI_soils_withNAs.csv")

```

##Complete Data
Getting together only the samples that have a complete dataset 
```{r complete data, echo=FALSE}
completeset = soils_data_all_wmeta_vis %>% filter(collected == TRUE) %>%
    select(campaign, kit_id, transect_location, loi_perc, bulk_density_g_cm3, gwc_perc, tn_perc, tc_perc, specific_conductance_us_cm, ph, gwc_perc, region, percent_clay, percent_sand, plant_litter, visible_minerals, visible_iron_ox, sample_weight_g) %>% select(!ends_with("_flag")) %>% na.omit()

saveRDS(completeset, "./soils_data_foralex.rds")
```

```{r code for alex, echo=FALSE}
#get and load needed packages
require(googlesheets4)
library(googlesheets4)

require(tidyverse)
library(tidyverse)

#Give access permissions. Put in your email address.
#options(gargle_oautho_email= "allison.myers-pigg@pnnl.gov")

Alex_OC = googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wLfwNtOP7hw2B102C9ZmvJtLfuQtvpqM2KTTTiFDmy4/edit#gid=0", sheet = 1)

#clean up dataset before merging. Needs the same column headers as the other dataframe for the column you are merging, and need consistent capitalization or not in the kit names. Also the OC values need to be averaged. 

Alex_OC_clean = Alex_OC %>%
  #this line fixes the column names to make them R friendly.
  dplyr::rename(oc_perc_rep1 = `OC % Run 1`,
                oc_per_rep2 = `OC % Run 2`,
                carbon_type = `Carbon Type`,
                mineral_type = `Mineral Type`) %>%
  #this line extracts the kit_id, transect location and creates the campaign name from the kit name into its own column(s)
    mutate(kit_id = stringr::str_extract(KIT_ID,"K[:digit:]{3}" ), #extract kit ID
           KIT_ID = tolower(KIT_ID), #this gives a consistent case for the sample names in the original kit ID column. This goes after we extract the uppercase K from KIT_ID so we don't have to change it back later.
           transect_location = stringr::str_extract(KIT_ID,"upland|transition|wetland|sediment" ), #extract transect location 
           campaign = "EC1") %>% #make campaign name. 
#remove columns we don't need anymore 
  select(-KIT_ID) %>%
  #get dataframe ready to take mean OC
  pivot_longer(cols = c(oc_perc_rep1, oc_per_rep2),
               names_to = 'repnumber') %>%
  rename(oc_per = value) %>%
  group_by(kit_id, transect_location, campaign, carbon_type, mineral_type) %>%
  dplyr::summarise(mean_oc_per = mean(oc_per, na.rm = TRUE), sd_oc_per = sd(oc_per, na.rm=TRUE)) 


#merge the two sheets. left_join will get you only the samples you have OC for: 

Alex_all_data <- left_join(Alex_OC_clean, soils_data_all_wmeta_vis, by=c("kit_id", "transect_location", "campaign"))

```
