---
title: "Exchange"
author: "Donnie Day"
date: "7/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#import
```{r}
df_cl<-read.csv("C:/Users/Donnie Day/Downloads/EC1_Metadata_CollectionLevel_070622_DD.csv")
df_kl<-read.csv("C:/Users/Donnie Day/Downloads/EC1_Metadata_KitLevel_070622_DD.csv")
```

```{r}
library(tidyverse)
```
##dataset_Collection Level 
#Grouping of macrophytes and algae to macrophytes
```{r}
df_cl[, 'water_macrophytes_algae'] <- case_when(
  df_cl[, 'water_macrophytes_algae'] == "None" ~ "None",
  df_cl[, 'water_macrophytes_algae'] == "Algae" ~ "Algae",
  df_cl[, 'water_macrophytes_algae'] == "Macrophytes" ~ "Macrophytes",
  df_cl[, 'water_macrophytes_algae'] == "Macrophytes, Algae" ~ "Macrophytes",
  df_cl[, 'water_macrophytes_algae'] == "NA" ~ "NA"
)
```

#Numeric classification of none, algae, macrophytes, and NA
```{r}
df_cl[, 'water_macrophytes_algae_num'] <- case_when(
  df_cl[, 'water_macrophytes_algae'] == "None" ~ 1,
  df_cl[, 'water_macrophytes_algae'] == "Algae" ~ 2,
  df_cl[, 'water_macrophytes_algae'] == "Macrophytes" ~ 3,
  df_cl[, 'water_macrophytes_algae'] == "NA" ~ NaN
)

```


#Numeric classification of H, O, A, and C horizons class wetland
```{r}
df_cl[, 'wt_Simple.Horizon.designation_num'] <- case_when(
  df_cl[, 'wt_Simple.Horizon.designation'] == "H" ~ 1,
  df_cl[, 'wt_Simple.Horizon.designation'] == "O" ~ 2,
  df_cl[, 'wt_Simple.Horizon.designation'] == "A" ~ 3,
  df_cl[, 'wt_Simple.Horizon.designation'] == "C" ~ 4,
  df_cl[, 'wt_Simple.Horizon.designation'] == "NA" ~ NaN,
  df_cl[, 'wt_Simple.Horizon.designation'] == "" ~ NaN
)
```

#Numeric classification of H, O, A, and C horizons class transition
```{r}
df_cl[, 'tr_Simple.Horizon.designation_num'] <- case_when(
  df_cl[, 'tr_Simple.Horizon.designation'] == "H" ~ 1,
  df_cl[, 'tr_Simple.Horizon.designation'] == "O" ~ 2,
  df_cl[, 'tr_Simple.Horizon.designation'] == "A" ~ 3,
  df_cl[, 'tr_Simple.Horizon.designation'] == "C" ~ 4,
  df_cl[, 'tr_Simple.Horizon.designation'] == "NA" ~ NaN,
  df_cl[, 'tr_Simple.Horizon.designation'] == "" ~ NaN
)
```

#Numeric classification of H, O, A, and C horizons class upland
```{r}
df_cl[, 'up_Simple.Horizon.designation_num'] <- case_when(
  df_cl[, 'up_Simple.Horizon.designation'] == "H" ~ 1,
  df_cl[, 'up_Simple.Horizon.designation'] == "O" ~ 2,
  df_cl[, 'up_Simple.Horizon.designation'] == "A" ~ 3,
  df_cl[, 'up_Simple.Horizon.designation'] == "C" ~ 4,
  df_cl[, 'up_Simple.Horizon.designation'] == "NA" ~ NaN,
  df_cl[, 'up_Simple.Horizon.designation'] == "" ~ NaN
)
```

##dataset_Kit Level 
#Numeric classification of weather_simplified Sunny, cloudy, Overcast, NA
```{r}
df_kl[, 'weather_simplified_num'] <- case_when(
  df_kl[, 'weather_simplified'] == "Sunny" ~ 1,
  df_kl[, 'weather_simplified'] == "Cloudy" ~ 2,
  df_kl[, 'weather_simplified'] == "Overcast" ~ 3,
  df_kl[, 'weather_simplified'] == "NA" ~ NaN,
  
)
```



