---
title: "TAI Geochemistry Analysis"
author: "AMP, KH, DD"
date: '2022-08-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load in Data
```{r Load Data, echo=FALSE, results='hide'}
library(tidyverse)
soilparty = readRDS("./soils_data_merged_withmeta.rds")

```

##Complete Data
Getting together only the samples that have a complete dataset 
```{r data, echo=FALSE}
completeset = soilparty %>% filter(collected == TRUE) %>%
    select(campaign, kit_id, transect_location, loi_perc, bulk_density_g_cm3, gwc_perc, tn_perc, tc_perc, specific_conductance_us_cm, ph, gwc_perc, region, percent_clay, percent_sand, plant_litter, visible_minerals, sample_weight_g) %>% select(!ends_with("_flag")) %>% na.omit()

saveRDS(completeset, "./soils_data_foralex.rds")
```


## Clean Data for multivariate stats
#I cant get this function to work as a list??
dplyr::select(!acidification)
dplyr::select(!date_run) %>%
dplyr::select(!root_mass) %>%
dplyr::select(!collection_date) %>%
dplyr::select(!site_name) %>%
dplyr::select(!state) %>%

dplyr::select(!samples_collected) %>%
dplyr::select(!weather_conditions) %>%
dplyr::select(!kit_notes) %>%

soils4NMDS = soils4NMDS %>% mutate(air_temperature_c = stringr::str_replace(air_temperature_c, "-[0-9]{2}", ""))
soils4NMDS$air_temperature_c = as.numeric(soils4NMDS$air_temperature_c)
soils4NMDS$barometric_pressure_inhg= as.numeric(soils4NMDS$barometric_pressure_inhg)


```{r clean for multivariate stats, echo=FALSE}

soils4multi = completeset %>% unite(uniqueid, c(campaign,kit_id,transect_location)) %>% 
  dplyr::select(!region) %>%
column_to_rownames(var="uniqueid")


# soils4multi$visible_minerals <- factor(soils4multi$visible_minerals, levels =c("None", "Mica", "Quartz"))
# soils4multi$visible_minerals <- as.integer(soils4multi$visible_minerals)
# 
# soils4multi$visible_iron_oxidation <- factor(soils4multi$visible_iron_oxidation, levels =c("No", "Yes"))
# soils4multi$visible_iron_oxidation <- as.integer(soils4multi$visible_iron_oxidation)
# 
# soils4multi$soil_horizon <- factor(soils4multi$soil_horizon, levels =c("Organic", "Mineral", "Organic, Mineral"))
# soils4multi$soil_horizon <- as.integer(soils4multi$soil_horizon)


```

This is a mess: 

```{r Playing, echo=FALSE}
library(vegan)
library(dplyr)
library(ggplot2)
library(ggfortify)
library("wesanderson")

# soilspca = prcomp(soils4multi, retx = TRUE, center = TRUE, scale. = TRUE)
# biplot(soilspca)

soilsnmds = metaMDS(soils4multi, distance="bray", 
          k = 3,
          maxit = 999, 
          trymax = 500,
          wascores = TRUE)

transpal = c("#056009","#8E7941","#021677")

regionloc <- completeset %>% select(transect_location, region)

env1 <- envfit(soilsnmds, regionloc, permu = 999)

ordiplot (soilsnmds, display = 'si', type = 'n')
points (soilsnmds, col = alpha(transpal, 0.5), pch = 19)
ordiellipse(soilsnmds, groups=completeset$transect_location, col=c("#056009","#8E7941","#021677"))
#plot(env1, p.max = 0.05, label=FALSE)


# soilsnmds %>% ordiplot(type='points') %>% points(col=c("#8E7941","#DCCD6A","#27220F")) %>% ordiellipse(groups=completeset$transect_location, col=c("#8E7941","#DCCD6A","#27220F"),label = TRUE)


#wetland = "#021677
#transition = "#056009"
#upland = "#8E7941"

# pal <- wes_palette("Darjeeling2",3, type = "discrete")
# 
# graphy <-autoplot(soilspca, data = completeset, shape= 'region', colour= 'transect_location',
#   size=2,
#   loadings = TRUE, loadings.colour = 'black',
#   loadings.label = TRUE, loadings.label.size = 2)+
#   scale_colour_manual(values = pal) + 
#   cowplot::theme_cowplot()
# 
# print(graphy)

```

```{r} 
# soils_distmat <- 
#   vegdist(soils4NMDS, method = "bray")
# 
# soils_distmat <- 
#   as.matrix(soils_distmat, labels = T)
# #write.csv(soils_distmat, "soils_distmat.csv")
# 
# Soils_NMDS <-
#   metaMDS(soils_distmat,
#           distance = "bray",
#           k = 3,
#           maxit = 999, 
#           trymax = 500,
#           wascores = TRUE)
# # Run 20 stress 0.0334468 
# # Procrustes: rmse 0.0009725123  max resid 0.002912856 
# goodness(Soils_NMDS)
# stressplot(Soils_NMDS)
# 
# #NMR_NMS_1 <-
# #  metaMDS(NMR_distmat,
# #          distance = "bray",
# #          k = 3,
# #          maxit = 999, 
#  #         trymax = 500,
#  #         wascores = TRUE)
# #Run 20 stress 0.03344632 
# # Procrustes: rmse 0.0002681262  max resid 0.001004477 
# #goodness(NMR_NMS_1)
# #stressplot(NMR_NMS_1)
```

